---
title: "Extracting NIBRS Data for the Network Analysis of Clearance Crime Rates Across the Country
"
author: "Mahlika George"
output: pdf_document
---
## Storing Variables 

To start the process of exporting the data into a formulated excel sheet, I first read in the data from the stored csv. The variables are than stored accordingly. 


```{r}
Data <- read.csv("~/Downloads/Sorted.csv", header = TRUE)

City <- Data$BH007

State <- Data$BH008

cities <- list()
#The following loop goes through all of the entries of cities listed in NIBRS and adds them to a new list only if they have not been added already. This is to remove duplicates.
for (i in City) {

    if (length(cities)==0) {

      cities <- append(cities,i)

    }

    else if (!i %in% cities) {

      cities <- append(cities,i)

    }

}

#The following code resaves city to only store the top 100 cities by population
cities <- head(cities,-(length(cities)-100))

#The following stores the index in where the last city first appears in NIBRS so we can find the associated state
#This loop finds where that city and state combo are last used in NIBRS so we can create a shortened data frame
index <- which(City %in% cities[100])[1]

s <- State[index]

for (i in which(City %in% cities[100])) {

  if(State[i] == s) {

    iny <- i

  }

  else{

    break

  }

}


#A new data fram is created for only the top 100 cities
new_Data <- Data[1:iny,]

```

## Calculating the Incidents and Rates

Next, I then created several lists to calculate and store the total number of incidents of certain crimes and all crimes. That data is calculated with a series of loops and conditionals. 


```{r}
#Several new variables are created for future storing

Counted <- list() #Total Incidents

sum <- list()

AA <- list()

AA_Clear <- list()

Murder <- list()

Murder_Clear <- list()

Robbery <- list()

Robbery_Clear <- list()

#This variable is used to iterate through the list as the starting index
index <- 1

#For numbers between 1 and the last index of the new data frame
for (i in 1:iny) {
  
#if the city has not yet been added to the list "Counted", and the sum is now 1, while every other variable is 0 for that index.
    if(!as.character(new_Data$BH007[i]) %in% Counted) {

    Counted[index] <- as.character(new_Data$BH007[i])

    Murder[index] <-0

    Murder_Clear[index] <- 0

    AA[index] <- 0

    AA_Clear[index] <- 0

    Robbery[index] <- 0

    Robbery_Clear[index] <-0

    sum[index] <- 1
    #if the offense code is 91, it is a murder. We initialize the murder count to 1 and stop checking. 
    if(new_Data$V20061[i]==91 | new_Data$V20062[i]==91 | new_Data$V20063[i] == 91) {

              Murder[index] <- 1
              
              
              #if there is an arrest, set the amount of cleared incidents to 1
               if(new_Data$V1011[i]>0) {

                Murder_Clear[index] <- 1

              }

            }

            #if the offense code is 131, it is an aggravated assault. We initialize aa to 1 and stop checking
              else if(new_Data$V20061[i]==131 | new_Data$V20062[i]==131 | new_Data$V20063[i] == 131) {

                AA[index] <- 1
                
                #if there is an arrest, set the amount of cleared incidents to 1
                 if(new_Data$V1011[i]>0) {

                  AA_Clear[index] <- 1

                }

              }
    
            #if the offense code is 120, it is a robbery. We initialize robbery to 1 and stop checking
            else if(new_Data$V20061[i]==120 | new_Data$V20062[i]==120 | new_Data$V20063[i] == 120 ){

                  Robbery[index] <- 1

                  if(new_Data$V1011[i]>0) {

                    Robbery_Clear[index] <- 1

                  }

              }
              
            #move to the next index
            index <- index + 1
  }
    #if that city has already been added to the list...        
  else {
    
    #add one to the current incident count
    sum[match(as.character(new_Data$BH007[i]),Counted)] <- 1 + as.numeric(sum[match(as.character(new_Data$BH007[i]),Counted)])
    
    #if its a murder, add one to the current murder count 
    if (new_Data$V20061[i]==91 | new_Data$V20062[i]==91 | new_Data$V20063[i] ==91) {

      Murder[match(as.character(new_Data$BH007[i]),Counted)] <- 1 + as.numeric(unlist(Murder[match(as.character(new_Data$BH007[i]),Counted)]))
      
      #if there is an arrest, add 1 to cleared incident
       if(new_Data$V1011[i]>0) {

        Murder_Clear[match(as.character(new_Data$BH007[i]),Counted)] <- 1 + as.numeric(unlist(Murder_Clear[match(as.character(new_Data$BH007[i]),Counted)]))
        
      }

    }

    #if its an aggravated assault, add one to the current count 
    else if (new_Data$V20061[i]==131 | new_Data$V20062[i]==131 | new_Data$V20063[i]==131) {

      AA[match(as.character(new_Data$BH007[i]),Counted)] <- 1 + as.numeric(unlist(AA[match(as.character(new_Data$BH007[i]),Counted)]))
      
      #if there is an arrest, add 1 to cleared incident
      if(new_Data$V1011[i]>0) {

        AA_Clear[match(as.character(new_Data$BH007[i]),Counted)] <- 1 + as.numeric(unlist(AA_Clear[match(as.character(new_Data$BH007[i]),Counted)]))

      }

    }
    
    #if its a robbery, add one to the current robbery count 
    else if (new_Data$V20061[i]==120 | new_Data$V20062[i]==120 | new_Data$V20063[i]==120) {

      Robbery[match(as.character(new_Data$BH007[i]),Counted)] <- 1 + as.numeric(unlist(Robbery[match(as.character(new_Data$BH007[i]),Counted)]))

      #if there is an arrest, add 1 to cleared incident
      if(new_Data$V1011[i]>0) {

        Robbery_Clear[match(as.character(new_Data$BH007[i]),Counted)] <- 1 + as.numeric(unlist(Robbery_Clear[match(as.character(new_Data$BH007[i]),Counted)]))

      }

    }

  }

}

#Store the final city list as the Counted list
final_city <- Counted

final_state <- list()

final_population <- list()

type_incidents <- list()

crime_rate <- list()

murder_perct <- list()

murder_clearance_rate <- list()

aa_perct <- list()

aa_clearance_rate <- list()

robbery_perct <- list()

robbery_clearance_rate <- list()

j <-1

#For every city in the Counted list, add there corresponding State from the data frame, in the same matching index but within a new list "final_state"
for (i in final_city) {

  final_state[j] <- as.character(new_Data$BH008[match(i,new_Data$BH007)])

  #adding the populations to a list based on the cities
  final_population[j] <- as.character(new_Data$BH019[match(i,new_Data$BH007)])
  
  #the specific type of incidents is the sum of the 3 interested crimes
  type_incidents[j] <- as.character(as.numeric(AA[j]) + as.numeric(Robbery[j]) + as.numeric(Murder[j]))
  
  #crime rate is type incidents divided by population times 100000
  crime_rate[j] <- (as.numeric(type_incidents[j])/as.numeric(final_population[j])) * 100000
  
  #percentage of murders
  murder_perct[j] <- as.numeric(Murder[j])/as.numeric(type_incidents) * 100
  
  murder_clearance_rate[j] <- (as.numeric(Murder_Clear[j])/as.numeric(Murder[j]))*100
  
  #percentage of aggravated assaults
  aa_perct[j] <- as.numeric(AA[j])/as.numeric(type_incidents) * 100
  
  aa_clearance_rate[j] <- (as.numeric(AA_Clear[j])/as.numeric(AA[j]))*100
  
  #percentage of robbery
  robbery_perct[j] <- as.numeric(Robbery[j])/as.numeric(type_incidents) * 100
  
  robbery_clearance_rate[j] <- (as.numeric(Robbery_Clear[j])/as.numeric(Robbery[j]))*100
  
  j <- j + 1
  }

```

Now that all of the values have been calculated, write it out to a new excel sheet


```{r}
#Download needed libraries
library("xlsx")

#Creating a new data frame of all of the new variables, and unlisting them so they are in the right format to do so
excel <- data.frame("Total Population" = matrix(unlist(final_population)), "City" = matrix(unlist(final_city)), "State" = matrix(unlist(final_state)), "Total Incidents by city" = matrix(unlist(sum)), "Total Interested Incidents by city" = matrix(unlist(type_incidents)), "Crime Rate" = matrix(unlist(crime_rate)), "Aggravated Assault Incidents" = matrix(unlist(AA)), "Aggravated Assault Percentage" = matrix(unlist(aa_perct)), "Aggravated Assault Incidents Cleared" = matrix(unlist(AA_Clear)),"Aggravated Assault Clearance Rate" = matrix(unlist(aa_clearance_rate)), "Murder/Nonnegligent Manslaughter Incidents" = matrix(unlist(Murder)), "Murder/Nonnegligent Manslaughter Percentage" = matrix(unlist(murder_perct)), "Murder/Nonnegligent Manslaughter Incidents Cleared" = matrix(unlist(Murder_Clear)),"Murder/Nonnegligent Manslaughter Clearance Rate" = matrix(unlist(murder_clearance_rate)), "Robbery Incidents" = matrix(unlist(Robbery)), "Robbery Percentage" = matrix(unlist(robbery_perct)), "Robbery Incidents Cleared" = matrix(unlist(Robbery_Clear)),"Robbery Clearance Rate" = matrix(unlist(robbery_clearance_rate)))


write.xlsx2(excel, file = "testing.xlsx", col.names = TRUE, row.names = TRUE)

```

